---
title: "R Notebook"
output: html_notebook
---

#Introduction
I will perform DEG analysis on microarray data from NCBI GEO series GSE99586 submitted by Miska and Sarkies. The NCBI site does not include a citation for any publication using this data, and I have not investigated that any further.

*I will be following the same workflow I used to investigate JU1580.*

I wanted to follow the same workflow I used to investigate JU1580, however, this data set was generated using a different type of microarray platform. Therefore I needed to download an install additional packages and alter certain parts of the analysis.

#Setup  
##Load Libraries
```{r eval=F}
#source("https://bioconductor.org/biocLite.R")
#biocLite("pd.elegene.1.0.st")
#biocLite("oligo")
library(gcrma)
library(limma)
library(affy)
library(annotate)
library(celegans.db)
library(GEOquery)
library(org.Ce.eg.db)
library(oligo)
library(oligoClasses)
library("pd.elegene.1.0.st", lib.loc="~/R/x86_64-pc-linux-gnu-library/3.5") #Specific for Ryan's laptop
```

##Unpack Data
I've download the data in an archival format, .tar, the first step is to unpack the data. It should contain a single cel file for each sample run.

```{r eval=F}
untar("GSE99586_RAW.tar", exdir = "GSE99586_data")
```

##Load cel files
Now that we have unpacked the files, we will import them to R and continue our analysis. Note all the cel files are compressed and we must uncompress them.
```{r eval=F}
cels <- list.files("GSE99586_data/", pattern = "CEL")
sapply(paste("GSE99586_data", cels, sep="/"), gunzip)
cels <- list.celfiles("GSE99586_data/", full.names = T)
cels
```

```{r}
rawData <- read.celfiles(cels)
rawData
```


##Organize data
Success, we have unpacked all the compressed cel files and now we will continue our analysis. The first function `ReadAffy()` is in the `affy` library we loaded earlier. It will read cel files and concatenate them into an affybatch object. The following function `rma()` will convert the affybatch object into an ExpressionSet object. These ExpressionSet objects can be manipulated by many existing R packages.
```{r}
eset <- oligo::rma(rawData)
head(exprs(eset))
boxplot(eset)
```

##Rename columns
Remove the long and cumbersome GSE names and .cel extensions. Additionally, R does not like to have numbers as the first character of a column. (I think you can if you specify that the number is a character string, but I haven't tested this.) In this case I make a simple substitution and swap the `6` for `"R"`. It won't matter though since we are not concerned with those samples.
```{r}
samplenames <- substring(colnames(eset),24, nchar(colnames(eset))-4)
samplenames <- gsub(6,"R",samplenames)
colnames(eset) <- samplenames
head(exprs(eset))
```

```{r}
par(mfrow=c(1,2))
plotMDS(eset)
plotMDS(eset, dim.plot = c(3,4))
```

#Data organization
##Match chip cells with c elegans genes
Turns out, after many hours of tinkering and googling, that the probe set IDs of this particular array are not included in the annotation dbs (like `celegans.db`). This `getNetAffx()` function is the best thing that I could find. In this case, it matches the probeID with the gene and protein information. I think this data is contained in the `eset` object, but I am not positive. I am also not positive on how to access the annotation information until later, when I call `topTable()`.
```{r}
annot <- getNetAffx(eset, type = "transcript")
featureData(eset) <- annot
```

Extract the first two characters for the groups. be sure to use `as.factor()` as the next step, design matrix, will require factors.
```{r}
group <- as.factor(substring(samplenames,1,2))
group
```

##Make Design matrix
```{r}
design <- model.matrix(~0+group)
colnames(design) <- levels(group)
design   
```

##Generate contrast matrix
There is only one comparisons I am interested in for this analysis. drh-1(-) vs N2 uninfected. 
```{r}
contrast_matrix <- makeContrasts(
  drh1.vs.n2 = DU - NU,
  drh1inf.vs.drh1.un = DI - DU,
  N2inf.vs.N2.un = NI - NU,
  levels = design
)
contrast_matrix
```

```{r}
fit <- lmFit(eset, design)
fit2 <- contrasts.fit(fit, contrast_matrix)
efit <- eBayes(fit2, trend = T)
results <- decideTests(efit)
summary(results)
```

```{r}
topTable(efit, coef=1, p.value = 0.05, n=Inf, sort.by = "M")
topTable(efit, coef=2, p.value = 0.05, n=Inf, sort.by = "M")
topTable(efit, coef=3, p.value = 0.05, n=Inf, sort.by = "M")
```

