---
title: "R Notebook"
output: html_notebook
---

#Introduction
I will perform DEG analysis on microarray data from NCBI GEO series GSE99586 submitted by Miska and Sarkies. The NCBI site does not include a citation for any publication using this data, and I have not investigated that any further.

*I will be following the same workflow I used to investigate JU1580.*

I wanted to follow the same workflow I used to investigate JU1580, however, this data set was generated using a different type of microarray. Therefore I needed to download an install additional packages

#Setup  
##Load Libraries
```{r}
source("https://bioconductor.org/biocLite.R")
#biocLite("pd.elegene.1.0.st")
biocLite("oligo")
library(gcrma)
library(limma)
library(affy)
library(annotate)
library(celegans.db)
library(GEOquery)
library(org.Ce.eg.db)
library(oligo)
library(oligoClasses)
library("pd.elegene.1.0.st", lib.loc="~/R/x86_64-pc-linux-gnu-library/3.5") #Specific for Ryan's laptop
```

##Unpack Data
I've download the data in an archival format, .tar, the first step is to unpack the data. It should contain a single cel file for each sample run.

```{r}
untar("GSE99586_RAW.tar", exdir = "GSE99586_data")
```

##Load cel files
Now that we have unpacked the files, we will import them to R and continue our analysis. Note all the cel files are compressed and we must uncompress them.
```{r}
cels <- list.files("GSE99586_data/", pattern = "CEL")
sapply(paste("GSE99586_data", cels, sep="/"), gunzip)
cels <- list.celfiles("GSE99586_data/", full.names = T)
cels
```

```{r}
rawData <- read.celfiles(cels)
rawData
```


##Organize data
Success, we have unpacked all the compressed cel files and now we will continue our analysis. The first function `ReadAffy()` is in the `affy` library we loaded earlier. It will read cel files and concatenate them into an affybatch object. The following function `rma()` will convert the affybatch object into an ExpressionSet object. These ExpressionSet objects can be manipulated by many existing R packages.
```{r}
eset <- backgroundCorrect(rawData)
head(exprs(eset))
boxplot(eset)
```

